---
title: "Modern Dive Chapter 2"
author: "Cameron Lyons"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(comment = "")
```

***

Read [Chapter 2](https://moderndive.com/2-getting-started.html) of _Modern Dive_. Work through the activities below and complete this accompanying Markdown document. Knit to HTML and submit your work as instructed. 

### 2.1 What are R and RStudio?

You will be using AsU's [RStudio server](https://mathr.math.appstate.edu) in this course, though you are welcome (and encouraged) to install R and RStudio to your own computer to use offline, as per the directions in 2.1.1 Installing R and RStudio. After reading and watching the videos below, which come from the DataCamp module [Working with the R Studio IDE (Part 1)](https://www.datacamp.com/courses/working-with-the-rstudio-ide-part-1), answer the questions. You might want to at least skim through the entire DataCamp module for additional practice.

* [Introduction](https://campus.datacamp.com/courses/working-with-the-rstudio-ide-part-1/orientation?ex=1)
* [RStudio's Panes](https://campus.datacamp.com/courses/working-with-the-rstudio-ide-part-1/orientation?ex=5) 
* [The Source Pane](https://campus.datacamp.com/courses/working-with-the-rstudio-ide-part-1/orientation?ex=9)
* [The View() Function](https://campus.datacamp.com/courses/working-with-the-rstudio-ide-part-1/orientation?ex=12) 
* [The Environment Pane](https://campus.datacamp.com/courses/working-with-the-rstudio-ide-part-1/orientation?ex=15)
* [The Files Pane](https://campus.datacamp.com/courses/working-with-the-rstudio-ide-part-1/orientation?ex=20)
* [The Help Pane](https://campus.datacamp.com/courses/working-with-the-rstudio-ide-part-1/orientation?ex=27)

1) Explain in your own words the difference between R and RStudio and how they work together. Include brief descriptions of the purpose and function of the `Source`, `Environment`, `Files`, and `Help` panes. 

ANSWER: R is the engine behind RStudio. R is seen as the engine to a car while RStudio is seen as the steering wheel. `Source` purpose is to interact with R through RStudio. `Environment` purpose is to show how data is represented in Source. `Files` purpose is to access different coding files that have been created. `Help` purpose is a search engine to help code.


2) How do you turn on R style diagnostics?

ANSWER: Tools + Code + Diagnostics + Show diagnostics for R


3) What is the PC keyboard shortcut to do each of the following?

a) clear the console?

ANSWER: ctrl + l


b) comment or uncomment lines of code?

ANSWER: ctrl + shift + c


c) reflow comments?

ANSWER: ctrl + shift + /


d) find a line of code?

ANSWER: ctrl + f


e) source with echo?

ANSWER: Ctrl + Shift + Enter


f) restart R session?

ANSWER: cctrl + shift + F10


g) insert an assignment operator (`<-`)?

ANSWER: alt + -


h) insert a pipe operator (`%>%`)?

ANSWER: ctrl + shift + M


i) insert a code chunk?

ANSWER: ctrl + alt + I


j) Keyboard Shortcut Quick Reference?

ANSWER: alt + shift + K


### 2.4 Explore your first dataset

Load all the packages you need.

```{r}
library(nycflights13)
library(tibble)
library(knitr)
```


### 2.4.1 nycflights13 package

The `nycflights13` package contains five datasets saved as data frames, with information about all domestic flights departing from from Newark Liberty International (EWR), John F. Kennedy International (JFK), or LaGuardia (LGA) airports in 2013.

* `flights`: information on all 336,776 flights
* `airlines`: translation between two letter IATA carrier codes and names (16 in total)
* `planes`: construction information about each of 3,322 planes used
* `weather`: hourly meteorological data (about 8705 observations) for each of the three NYC airports
* `airports`: airport names and locations


### 2.4.2 flights data frame

Enter `flights` in the code chunk below to get your first look at the data frame, then answer the questions.

```{r}
nycflights13::flights
```

How many observations are there in the data frame? 

ANSWER: 336,776

How many variables are there in the data frame? 

ANSWER: 19


### 2.4.3 Exploring data frames

1. `View()`

Run `View(flights)` in the RStudio Console and explore the `flights` data frame in the pop-up viewer that appears in the Source pane. What does any ONE row in this flights dataset refer to? 

ANSWER: A single flight to and from a certain destination.


2. `glimpse()`

Use `glimpse()` in the R code chunk below to explore the `flights` data frame, then answer the questions.

```{r}
glimpse(nycflights13::flights)
```

What are some examples in this dataset of categorical variables? What makes them different than quantitative variables? 

ANSWER: One example of a categorical variable is the distance variable. An example a a quantitative variable is departure time. 

What does int, dbl, and chr mean in the output above? 

ANSWER: integer, double, and character


3. `kable()

Enter `airlines` in the code chunk below to get your first look at the data frame, then use `kable()` further explore `airlines`.

```{r}
kable(airlines)
```


4. `$` operator

Display `airlines` and then use the `$` operator to extract only the `name` variable and return it as a vector.

```{r}
airlines$name
```


### 2.4.4 Help files

Use `?flights` to access the Help file for the `flights` data frame. According to the Help file, what does each column (variable) in the data frame represent?

* `year`: Date of departure

* `month`: Date of departure

* `day`: Date of departure 

* `dep_time`: Departure time

* `arr_time`: Arrival time

* `sched_dep_time`: scheduled departure time

* `sched_arr_time`: scheduled arrival time

* `dep_delay`: departure delay

* `arr_delay`: arrival delay

* `hour`: time of scheduled departure

* `minute`: time of scheduled departure

* `carrier`: different airlines

* `tailnum`: plane tail number

* `flight`: flight number

* `origin`: origin

* `dest`: destination

* `air_time`: amount of time spent in air

* `distance`: distance between airports, in miles

* `time_hour`: scheduled date and hour of the flight POSIXct date. Along with origin, can be used to join flights data to weather data.


***
END!