---
title: "R for Data Science CH03--CH05"
author: "Cameron Lyons"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(comment = "")
```

***

Read [Chapter 3](https://r4ds.had.co.nz/data-visualisation.html), [Chapter 4](https://r4ds.had.co.nz/workflow-basics.html), and [Chapter 5](https://r4ds.had.co.nz/transform.html) of _R for Data Science_. Work through the activities below and complete this accompanying Markdown document. Knit to HTML and submit your work as instructed. 

```{r, message = FALSE}
# load all packages here
```


### 3.5.1 Exercises

1) What happens if you facet on a continuous variable?

ANSWER: 


2) Read `?facet_wrap`. What does `nrow` do? What does `ncol` do? What other options control the layout of the individual panels? Why doesn’t `facet_grid()` have `nrow` and `ncol` arguments? Read the documentation [here](https://ggplot2.tidyverse.org/reference/facet_wrap.html) and [here](https://ggplot2.tidyverse.org/reference/facet_grid.html).

ANSWER:


### 3.6.1 Exercises

6) Recreate the R code necessary to generate the following graphs (see textbook).

```{r}
# show your code here (plot1)
```

```{r}
# show your code here (plot2)
```

```{r}
# show your code here (plot3)
```

```{r}
# show your code here (plot4)
```

```{r}
# show your code here (plot5)
```

```{r}
# show your code here (plot6)
```


### 3.9.1 Exercises

2) What does `labs()` do? Read the [documentation](https://ggplot2.tidyverse.org/reference/labs.html).

ANSWER:


3) What’s the difference between `coord_quickmap()` and `coord_map()`? Read the [documentation](https://ggplot2.tidyverse.org/reference/coord_map.html).

ANSWER:


4) What does the plot (see textbook) tell you about the relationship between city and highway mpg? Why is `coord_fixed()` important for this plot? What does `geom_abline()` do on this plot? Read the documentation [here](https://ggplot2.tidyverse.org/reference/coord_fixed.html) and [here](https://ggplot2.tidyverse.org/reference/geom_abline.html).

ANSWER:


### 5.2.4 Exercises

1) Find all flights that...

a) Had an arrival delay of two or more hours

```{r}
# show your code here
```


b) Flew to Houston (IAH or HOU)

```{r}
# show your code here
```


c) Were operated by United, American, or Delta

```{r}
# show your code here
```


d) Departed in summer (July, August, and September)

```{r}
# show your code here
```


e) Arrived more than two hours late, but didn’t leave late

```{r}
# show your code here
```


f) Were delayed by at least an hour, but made up over 30 minutes in flight

```{r}
# show your code here
```

g) Departed between midnight and 6am (inclusive)

```{r}
# show your code here
```


3) How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r}
# show your code here
```


### 5.3.1 Exercises

1) How could you use arrange() to sort all missing values to the start? (Hint: use `is.na()`).

```{r}
# show your code here
```


2) Sort flights to find the most delayed flights. Find the flights that left earliest.

```{r}
# show your code here
```


3) Sort flights to find the fastest flights.

```{r}
# show your code here
```


4) Which flights travelled the longest? Which travelled the shortest?

```{r}
# show your code here
```


### 5.5.2 Exercises

1) Currently `dep_time` and `sched_dep_time` are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

```{r}
# show your code here
```


2) Compare `dep_time`, `sched_dep_time`, and `dep_delay`. How would you expect those three numbers to be related? Do they meet this expectation?

```{r}
# show your code here
```

ANSWER:


### 5.7.1 Exercises

2) Which plane (tailnum) has the worst on-time record?

```{r}
# show your code here
```

ANSWER: 


3) What time of day should you fly if you want to avoid delays as much as possible?

```{r}
# show your code here
```

ANSWER:


***

```{r}
sessionInfo()
```
